\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-11-18 chapters/wcble.tex
Textcites

Most of the algorithms above can be adapted to work on multiple \ac{erp} components
by carefully selecting peaks or pre-determined regions of interest.
\textcite{Hardiansyah2020} uses an \ac{snr} boosting method, applied to a specific
time-window for each \ac{erp} component to extract a latency per component.
Methods based on spatial decomposition filters like \ac{pca}, \ac{ica}, or XDAWN can be leveraged
to utilize multiple of their components as filters, resulting in multiple time
series where peak picking or Woody's algorithm can be applied. In general,
however, these algorithms lack an integrated approach to deal with multiple
\ac{erp} components. A clear evaluation of the performance of these
methods in multi-component settings is lacking.

@@

\Ac{cble}, summarized in \cref{sec:wcble/app/cble-alg}, works by training a
\textit{first-stage} classifier $\mathcal{C}(\theta,f)$
defined within a time period $[s_1:s_2]$ with a set of parameters $\theta$ and a
decision function $f(\mat{X}[:,s_1:s_2],\theta) \rightarrow
y$ outputting a classification score $y\in\mathbb{R}$ for a given epoch $\mat{X}$,
such that
\begin{equation}
  \theta = \mathrm{train}_\mathcal{C}(\{\mat{X}_n^\mathrm{train}[:,s_1:s_2]\}_{n=1}^N,\mathbf{l}^\mathrm{train})
  \label{eq:cble-train}
\end{equation}
Then,
$f$ can be applied to all (possibly overlapping) slices of length $s_2-s_1$ of
an epoch $\mat{X}$, resulting in a vector of score values
$\mathbf{y}=[y_1\,\ldots\,y_R]^T \in\mathbb{R}^R$ such that
\begin{equation}
  y_s = f(\mat{X}[:,s:s+(s_2-s_1)],\theta)\quad \forall s\in 1,\ldots,R
	\label{eq:score}
\end{equation}
with $R = S-(s_2-s_1)$.
To leverage \ac{cble} for \ac{erp} classification, the score vectors $\mathbf{y}$ can be
arranged in matrices $\mat{Y}^\mathrm{train}\in\mathbb{R}^{N\times R}$ and $\mat{Y}^\mathrm{test}\in\mathbb{R}^{M\times R}$.
These can be further classified by training a \textit{second-stage} classifier on
$\mat{Y}^\mathrm{train}$ and class labels $\mathbf{l}^\mathrm{train}$.
However, the resulting score-over-time vectors per epoch still suffer from jitter.
For classification, we follow the approach of \cite{Mowla2017}, using a
maximum-level hierarchical Daubechies-4 wavelet transform to reduce
dimensionality before classification with the second-stage classifier.
In the \ac{cble}-decoder, it is this wavelet transform that decreases the sensitivity
to latency differences, actively compensating for \ac{erp} latency jitter.

@@

However, including these latency features gives rise to the following issue when
classifying unseen data.
The \ac{cble} latency estimate is defined only for target epochs as
$s_\mathrm{target}=\argmax_s{y_s}$.
This is the point in time where the target class reaches the largest separation
from the background noise and non-target class, indicating the target \ac{erp} is
most likely to occur here.
However, in a classifier test phase, it is not known a priori whether an
unseen epoch is a target or a non-target epoch.
This problem is solved by defining an estimated latency per class
$s_\mathrm{target}$ and $s_\mathrm{non-target}$ for every epoch,
regardless of its actual class.
The estimated class latencies can then be used as features for training and
testing the second-stage classifier.
This way, the latencies of the testing data can be presented to the second-stage
classifier without knowledge of the testing data class labels, making them
useful in decoding.

>>>>

Most of the algorithms above can be adapted to work on multiple \ac{erp} components
by carefully selecting peaks or pre-determined regions of interest.
\textcite{Hardiansyah2020} use an \ac{snr} boosting method, applied to a specific
time-window for each \ac{erp} component to extract a latency per component.
Methods based on spatial decomposition filters like \ac{pca}, \ac{ica}, or XDAWN can be leveraged
to utilize multiple of their components as filters, resulting in multiple time
series where peak picking or Woody's algorithm can be applied. In general,
however, these algorithms lack an integrated approach to deal with multiple
\ac{erp} components. A clear evaluation of the performance of these
methods in multi-component settings is lacking.

@@

\Ac{cble}, summarized in \cref{sec:wcble/app/cble-alg}, works by training a
\textit{first-stage} classifier $\mathcal{C}(\theta,f)$
defined within a time period $[s_1:s_2]$ with a set of parameters $\theta$ and a
decision function $f(\mat{X}[:,s_1:s_2],\theta) \rightarrow
y$ outputting a classification score $y\in\mathbb{R}$ for a given epoch $\mat{X}$,
such that
\begin{equation}
  \theta = \mathrm{train}_\mathcal{C}(\{\mat{X}_n^\mathrm{train}[:,s_1:s_2]\}_{n=1}^N,\mathbf{l}^\mathrm{train})
  \label{eq:cble-train}
\end{equation}
Then,
$f$ can be applied to all (possibly overlapping) slices of length $s_2-s_1$ of
an epoch $\mat{X}$, resulting in a vector of score values
$\mathbf{y}=[y_1\,\ldots\,y_R]^T \in\mathbb{R}^R$ such that
\begin{equation}
  y_s = f(\mat{X}[:,s:s+(s_2-s_1)],\theta)\quad \forall s\in 1,\ldots,R
	\label{eq:score}
\end{equation}
with $R = S-(s_2-s_1)$.
To leverage \ac{cble} for \ac{erp} classification, the score vectors $\mathbf{y}$ can be
arranged in matrices $\mat{Y}^\mathrm{train}\in\mathbb{R}^{N\times R}$ and $\mat{Y}^\mathrm{test}\in\mathbb{R}^{M\times R}$.
These can be further classified by training a \textit{second-stage} classifier on
$\mat{Y}^\mathrm{train}$ and class labels $\mathbf{l}^\mathrm{train}$.
However, the resulting score-over-time vectors per epoch still suffer from jitter.
For classification, we follow the approach of \textcite{Mowla2017}, using a
maximum-level hierarchical Daubechies-4 wavelet transform to reduce
dimensionality before classification with the second-stage classifier.
In the \ac{cble}-decoder, it is this wavelet transform that decreases the sensitivity
to latency differences, actively compensating for \ac{erp} latency jitter.

@@

However, including these latency features gives rise to the following issue when
classifying unseen data.
The \ac{cble} latency estimate is defined only for target epochs as
$\textstyle{s_\mathrm{target}=\argmax_s{y_s}}$.
This is the point in time where the target class reaches the largest separation
from the background noise and non-target class, indicating the target \ac{erp} is
most likely to occur here.
However, in a classifier test phase, it is not known a priori whether an
unseen epoch is a target or a non-target epoch.
This problem is solved by defining an estimated latency per class
$s_\mathrm{target}$ and $s_\mathrm{non-target}$ for every epoch,
regardless of its actual class.
The estimated class latencies can then be used as features for training and
testing the second-stage classifier.
This way, the latencies of the testing data can be presented to the second-stage
classifier without knowledge of the testing data class labels, making them
useful in decoding.
